Robustly decorrelating errors with mixed quantum gates

### 2019-09-19 decorrelation/mixed_unitaries.tex
Text and plot updates

% ==============================================================================
% Section: Constructing Useful mixed quantum gates
% ==============================================================================
\section{Constructing mixed quantum gates}
\label{sec:mixed_unitary_processes}
\noindent In this section, we present a methodology for constructing mixed quantum channels, formalizing the intuitive approach discussed above. As mentioned, our method requires two steps. The first is a control synthesis step, in which we construct an ensemble of gate implementations. Campbell and Hastings draw their ensembles from various different gate compilations, but for this work we use utilize quantum optimal control theory. Many standard control-generation algorithms, such as GRAPE \cite{Khaneja2005}, take a random initial guess for the control and iteratively improve it to yield a gate that well-approximates a target. By seeding such algorithms with many initial guesses, one can quickly construct a large ensemble of approximate quantum gates, each with a different error channel. We discuss this approach in some detail in Sec.~\ref{sec:numerical_results}. \kcy{Why should the gates we generate have any error at all? They shouldn't if we know the model. Discuss this here.}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{vectorspace3.png}
  \caption{\textbf{Geometric condition for robustness:}   A target unitary gate can be implemented a number of ways, each with a different effective error generator, $\genmat_i$. a) The error generators are shown as elements of a vector space (solid, black arrows). Also shown are their derivatives $\genmat_{i,k}$ (dashed, grey arrows) with respect to a model parameter, $\delta_k$. As this parameter drifts, the generators may no longer cancel. b) To construct a robust, generator-exact MQG, there must exist a set of weights, $\{\weight_i\}$, so that $\sum_i \weight_i \genmat_i = \sum_i \weight_i \genmat_{i,k} = 0$. This is equivalent to demanding that the origin lie in the convex hull $\mathsf{conv}(\{\genmat_i \bigoplus_k \genmat_{i,k} \}_i)$. }
  \label{fig:vectorspace2}
\end{figure}

@@

If we do not expect to have a space of generator-exact solutions available, then the constraint above will never be satisfied and the minimization will fail. This could occur if there are simply not enough controls to satisfy the generator-exact criterion, or if the error generators have stochastic components. In such a case, we can choose a weighting parameter, $\eta$, to explicitly balance the two competing objectives:

\begin{equation}\label{eq:agi_minimization_weighted}
  \begin{split}
    &\underset{\weight_i\geq0, |\weight|_1=1}{\textbf{minimize}: } ||\mathbf{L}\cdot\vec\weight||_2  + \eta \sum_i \weight_i\; \AGI(\errmat_i)
  \end{split}
\end{equation}

Both optimizations are convex and efficiently solvable.

@@

\subsection{Experimental implementation } % (fold)
\label{sub:experimental}
\noindent We implemented our methods using the 19Q-Acorn superconducting transmon processor at Rigetti Computing. Comprehensive device characterization for this processor can be found in \cite{1712.05771}. While this device consisted of 20 qubits with fixed couplings, we used only qubit \#8 for our demonstration. Controls were derived from a single, 10-sample, 50ns Gaussian $X_{\pi/2}$ pulse calibrated by iteratively amplifying and correcting over- and under-rotation errors via repeated application. Four intentionally miscalibrated Gaussian pulses, labeled $\mathsf{Pulse1}$ through $\mathsf{Pulse4}$, were derived from this waveform by scaling the amplitude of the calibrated $X_{\pi/2}$ pulse by $106.4\%$,  $103.9\%$, $93.7\%$ and $91.2\%$. The gates therefore experience unitary over/under rotations, which are coherent errors. These gates are then used as in Sec.~\ref{sub:off_diagonals} to construct a Pauli-exact MQG, which samples from the miscalibrated waveforms with probabilities, \kcy{What are they?}\amp{I'll have to go hunting, I didn't save these}, respectively.

We characterize the performance of the MQG using a series of randomized benchmarking (RB) experiments\cite{Magesan2011}: one for each of the miscalibrated pulses, one using the calibrated pulse, and using the MQG. In each case, the Clifford operations were decomposed into $X_{\pi/2}$ gates, which were implemented directly using the relevant waveforms, and $Y_{\pi/2}$ gates, which were implemented by phase-shifting the relevant waveform $\pi/2$ radians.  An RB experiment consisted of $10$ sequences sampled at each of lengths $\{2, 4, 8, 16, 32, 64\}$, and each sequence was repeated $1000$ times. We emphasize that, when benchmarking the MQG, each sequence repetition used a new, randomly sampled sequence of waveforms consistent with its definition.  Results of the standard RB analysis are indicated in Table.~\eqref{tabl:rb}. The MQG is seen to perform nearly as well as the calibrated pulse, and better than any of the constituent pulses individually.

Figure \ref{fig:rb} illustrates the distribution of survival probabilities at sequence length $L=64$ for each of the experiments.

\setlength{\tabcolsep}{0.5em} 
{\renewcommand{\arraystretch}{1.2}
% \newcolumntype{d}[1]{D{.}{.}{#1}}
\begin{table}[h]
	\label{tabl:rb}
	\centering
	\begin{tabular}{@{}ccc@{}}
		\hline
		Pulse name & Calibration error & RB error rate\\
		\hline
		$\mathsf{Pulse1}$ 	& 6.4\% 	& 1.1\% \\
		$\mathsf{Pulse2}$ 	& 3.9\% 	& 0.9\% \\
		$\mathsf{Pulse3}$ 	& --\,6.3\% 	& 1.1\% \\
		$\mathsf{Pulse4}$ 	& --\,8.8\% 	& 1.5\% \\
		Calibrated 			& $\cdot$  	& 0.7\% \\
		MQG 				& $\cdot$ 	& 0.8\% \\
		\hline
	\end{tabular}
	\caption{\textbf{Randomized benchmarking results:} The MQG outperforms all of the individual miscalibrated gates of which it is composed, achieving an RB number nearly as low as the carefully calibrated pulse. }
\end{table}

By minimizing the off-diagonal elements of the process matrix, we expect the resulting MQG to display suppressed coherent error. To see that this is in fact the case, we can inspect the variance of the $RB$ survival probabilities for the MQG relative to the miscalibrated pulses. As discussed in \cite{Ball2016}, coherent errors will tend to broaden the distribution of RB survival probabilities over sequences at each length, generally manifesting as a long-tailed gamma distribution. Stochastic noise, such as depolarizing noise, will yeilds relatively narrow, Gaussian distributed success probabilities. In Fig.~\ref{fig:rb}, we plot the distribution of survival probabilities at sequence length $64$ for each of the benchmarked gate sets. We see that the intentionally miscalibrated controls in our RB experiment have long tails, consistent with coherent noise, while the calibrated and randomized implementations are both significantly less dispersed, consistent with stochastic errors. This experiment therefore provides compelling evidence that the MQG suffers considerably less coherent error than any of the miscalibrated gates from which it was constructed.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{rb_data.pdf}
  \caption{\textbf{MQGs improve randomized benchmarking:}. The first four boxes result from using each of four different implementations of the ${\pi/2}$ rotations. The coherent noise present in these implementations leads to large variance of the survival probability over sequences. The fifth (dark blue) box illustrates the survival probability using a highly-tuned gate implementation. It displays improved average survival probability as well as reduced variance. The final box (dark red) illustrates the distribution over survival probabilities for a randomized MQG composed of $\mathsf{Pulse1}$ through $\mathsf{Pulse4}$. It performs comparably to the highly-calibrated implementation in both average survival probability and variance over random sequences. The reduced variance of the MQG is a tell-tale sign of reduced coherent error in the effective channel. }
  \label{fig:rb}
\end{figure}

We note that, while this example is somewhat contrived (the calibrated gate is clearly the best), it nonetheless succeeds in demonstrating that the MQGs are capable of outperforming the individual gates of which they are composed. In future work, we hope to apply this technique to two-qubit gates, which are harder to tune up in general and are more likely to possess lingering coherent errors \needcite.

@@

% ==============================================================================
% Section: Numerical Results
% ==============================================================================
\subsection{Numerical Implementation}
\label{sec:numerical_results}
\noindent In the following numerical results, we use the methods in Section \ref{sec:mixed_unitary_processes} to build MQGs, and then analyze their performance for a range of Hamiltonian parameters. More specifically, we will construct a family of controls with some degree of inherent robustness to drift, and leverage the numerical techniques introduced to assign each control a probability. When drawn from with these probabilities, the controls will form an MQG with enhanced robustness to drift on the control parameters. We consider the following dimensionless model for a single tunable qubit: 
\begin{equation}\label{eq:1Qham}
  H(\delta, \epsilon, t) = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
where $\epsilon$ corresponds to fluctuations in qubit frequency and $\delta$ corresponds to fluctuations in the control field.

@@

In our two-qubit example we consider the following model for two tunable qubits coupled by a resonant exchange interaction, similar to that in \cite{McKay2016}:
\begin{equation} \label{eq:2Qham}
\begin{split}
H(\vec{\delta}, \vec{\epsilon}, t) = &\sum_{j=1}^2(\epsilon_j\sigma_z^j + (1 + \delta_j)(c_x^j(t)\sigma_x^j + c_y^j(t)\sigma_y^j)) \\
&+ \frac{1}{10}(XX + YY)
\end{split}
\end{equation}

>>>>

% ==============================================================================
% Section: Constructing Useful mixed quantum gates
% ==============================================================================
\section{Constructing mixed quantum gates}
\label{sec:mixed_unitary_processes}
\noindent We now present a methodology for constructing mixed quantum channels, formalizing the intuitive approach discussed above with a series of explicit convex programs. As mentioned, our method requires two steps. The first is a control synthesis step, in which we construct an ensemble of gate implementations. Campbell and Hastings draw their ensembles from various different gate compilations, but for this work we use utilize quantum optimal control theory. Many standard control-generation algorithms, such as GRAPE \cite{Khaneja2005}, take a random initial guess for the control and iteratively improve it to yield a gate that well-approximates a target. By seeding such algorithms with many initial guesses, one can quickly construct a large ensemble of approximate quantum gates, each with a different error channel. We discuss this approach in some detail in Sec.~\ref{sec:numerical_results}.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{vectorspace3.png}
  \caption{\textbf{Geometric condition for robustness:}   A target unitary gate can be implemented a number of ways, each with a different effective error generator, $\genmat_i$. a) The error generators are shown as elements of a vector space (solid, black arrows). Also shown are their derivatives $\genmat_{i,k}$ (dashed, grey arrows) with respect to a model parameter, $\delta_k$. As this parameter drifts, the generators may no longer cancel. b) To construct a robust, generator-exact MQG, there must exist a set of weights, $\{\weight_i\}$, so that $\sum_i \weight_i \genmat_i = \sum_i \weight_i \genmat_{i,k} = 0$. This is equivalent to demanding that the origin lie in the convex hull $\mathsf{conv}(\{\genmat_i \bigoplus_k \genmat_{i,k} \}_i)$. The gradient arrows indicate that this space is higher-dimensional. }
  \label{fig:vectorspace2}
\end{figure}

@@

If we do not expect to have a space of generator-exact solutions available, then the constraint above will never be satisfied and the minimization will fail. This could occur if there are simply not enough controls to satisfy the generator-exact criterion, or if the error generators have stochastic components. In such a case, we can choose a weighting parameter, $\eta$, to explicitly balance the two competing objectives:
\begin{equation}\label{eq:agi_minimization_weighted}
  \begin{split}
    &\underset{\weight_i\geq0, |\weight|_1=1}{\textbf{minimize}: } ||\mathbf{L}\cdot\vec\weight||_2  + \eta \sum_i \weight_i\; \AGI(\errmat_i)
  \end{split}
\end{equation}
Both optimizations are convex and efficiently solvable.

@@

\subsection{Experimental implementation } % (fold)
\label{sub:experimental}
\noindent We implemented our methods using the 19Q-Acorn superconducting transmon processor at Rigetti Computing. Comprehensive characterization of this device can be found in \cite{1712.05771}. While this device consisted of 20 qubits with fixed couplings, we used only qubit \#8 for our demonstration. Controls were derived from a single, 10-sample, 50ns Gaussian $X_{\pi/2}$ pulse calibrated via Rabi oscillation. Four intentionally miscalibrated Gaussian pulses, labeled $\mathsf{Pulse1}$ through $\mathsf{Pulse4}$, were derived by scaling the amplitude of the calibrated $X_{\pi/2}$ pulse by a factor $S \in \{1.064, 1.039,0.937, 0.912\}$, respectively. We model their effect on the qubit state as perfectly unitary rotations, $X_{A \pi/2} = \exp{(-i A \pi/2 \sigma_x)}$. These gates are then used as in Sec.~\ref{sub:off_diagonals} to construct a Pauli-exact MQG, which samples from the miscalibrated waveforms with probabilities, \kcy{What are they?}\amp{I'll have to go hunting, I didn't save these}, respectively.

We characterize the performance of the MQG using a series of randomized benchmarking (RB) experiments\cite{Magesan2011}: one for each of the miscalibrated pulses, one using the calibrated pulse, and using the MQG. In each case, the Clifford operations were decomposed into $X_{\pi/2}$ gates, which were implemented directly using the relevant waveforms, and $Y_{\pi/2}$ gates, which were implemented by phase-shifting the relevant waveform $\pi/2$ radians.  An RB experiment consisted of sampling and running $10$ random Clifford sequences at each of lengths $L\in\{2, 4, 8, 16, 32, 64\}$. Each sequence was repeated $1000$ times. We emphasize that, when benchmarking the MQG, each sequence repetition used a new, randomly sampled sequence of waveforms consistent with its definition.  Results of the standard RB analysis are discussed in Table.~\eqref{tabl:rb}. The MQG is seen to perform nearly as well as the calibrated pulse, and better than any of the constituent pulses individually.

\setlength{\tabcolsep}{0.5em} 
{\renewcommand{\arraystretch}{1.2}
% \newcolumntype{d}[1]{D{.}{.}{#1}}
\begin{table}[h]
	\label{tabl:rb}
	\centering
	\begin{tabular}{@{}ccc@{}}
		\hline
		Pulse name & Rotation error & RB error rate\\
		\hline
		$\mathsf{Pulse1}$ 	& 6.4\% 	& 1.1\% \\
		$\mathsf{Pulse2}$ 	& 3.9\% 	& 0.9\% \\
		$\mathsf{Pulse3}$ 	& --\,6.3\% 	& 1.1\% \\
		$\mathsf{Pulse4}$ 	& --\,8.8\% 	& 1.5\% \\
		Calibrated 			& $\cdot$  	& 0.7\% \\
		MQG 				& $\cdot$ 	& 0.8\% \\
		\hline
	\end{tabular}
	\caption{\textbf{Randomized benchmarking results:} The MQG outperforms all of the individual miscalibrated gates of which it is composed, achieving an RB number nearly as low as the carefully calibrated pulse. }
\end{table}

By minimizing the off-diagonal elements of the process matrix, we expect the resulting MQG to display suppressed coherent error. To see that this is in fact the case, we can inspect the variance of the $RB$ survival probabilities for the MQG relative to the miscalibrated pulses. As discussed in \cite{Ball2016}, coherent errors will tend to broaden the distribution of RB survival probabilities over sequences at each length, generally manifesting as a long-tailed gamma distribution. Stochastic noise, such as depolarizing noise, will yield comparatively narrow, Gaussian-distributed success probabilities. In Fig.~\ref{fig:rb}, we plot the experimentally observed distribution of survival probabilities at sequence length $64$ for each of the benchmarked gate sets. We see that the intentionally miscalibrated controls in our RB experiment have long tails, consistent with coherent noise, while the calibrated and randomized implementations are both significantly less dispersed, consistent with stochastic errors. This experiment therefore provides compelling evidence that the MQG suffers considerably less coherent error than any of the miscalibrated gates from which it was constructed.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{rb_data.pdf}
  \caption{\textbf{Evidence of reduced coherent error:}. Shown here is the distribution of randomized benchmarking survival probabilities at $L=64$ for each of six native gate implementations. The first four boxes (light blue) result four intentionally miscalibrated ${\pi/2}$ rotations. The coherent noise present in these implementations leads to large variance in the survival probability over sequences. The fifth (dark blue) box illustrates the survival probability using a highly-tuned gate implementation. It displays improved average survival probability as well as reduced variance. The final box (dark red) illustrates the distribution for a randomized MQG composed of $\mathsf{Pulse1}$ through $\mathsf{Pulse4}$. It performs comparably to the highly-calibrated implementation in both average survival probability and variance over random sequences. The reduced variance of the MQG is a tell-tale sign of reduced coherent error in the effective channel \kcy{cite Biercuk's paper}. }
  \label{fig:rb}
\end{figure}

We note that, while this example is somewhat contrived (the calibrated gate is clearly the best), it nonetheless succeeds in demonstrating that the MQGs are capable of outperforming the individual gates of which they are composed. In future work, we hope to apply this technique to two-qubit gates, which are significantly more difficult to tune up in general and are more likely to possess lingering coherent errors. Tomography or modeling can then be used to construct the process matrix estimates required to build a mixed gate.

@@

% ==============================================================================
% Section: Numerical Results
% ==============================================================================
\subsection{Numerical implementations}
\label{sec:numerical_results}
\noindent In the following numerical results, we use the various methods in Section \ref{sec:mixed_unitary_processes} to build a range of one- and two-qubit MQGs for model systems. We analyze the resulting mixed gates for a range of Hamiltonian parameters, demonstrating both a reduction in diamond distance and an improved robustness to model uncertainty.

% ------------------------------------------------------------------------------
% Subsection: Single-qubit example
% ------------------------------------------------------------------------------
\subsection{Single-qubit example} 
\label{sub:one_qubit}
We consider the following dimensionless model for a single tunable qubit: 
\begin{equation}\label{eq:1Qham}
  H(\delta, \epsilon, t) = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
where $\epsilon$ corresponds to fluctuations in qubit frequency and $\delta$ corresponds to fluctuations in the control field.

@@

% ------------------------------------------------------------------------------
% Subsection: Single-qubit example
% ------------------------------------------------------------------------------
\subsection{Two-qubit example} 
\label{sub:two_qubit}
In our two-qubit example we consider the following model for two tunable qubits coupled by a resonant exchange interaction, similar to that in \cite{McKay2016}:
\begin{equation} \label{eq:2Qham}
\begin{split}
H(\vec{\delta}, \vec{\epsilon}, t) = &\sum_{j=1}^2(\epsilon_j\sigma_z^j + (1 + \delta_j)(c_x^j(t)\sigma_x^j + c_y^j(t)\sigma_y^j)) \\
&+ \frac{1}{10}(XX + YY)
\end{split}
\end{equation}
